Supplementary Figures

### 2024-09-03 manuscript_plots.Rmd
updates for review

## Figure 5
```{r model_type-site}
n_model_type <- temperature_skill |> 
  distinct(model_id) |> 
  left_join(model_meta) |>
  group_by(model_type) |> 
  summarise(n = n())  |> 
  mutate(site_id = 'TOOK', mean_skill = -3.25) |> 
  filter(model_type != 'null')

>>>>

## Figure 5
```{r model_type-site}
n_model_type <- temperature_skill |> 
  distinct(model_id) |> 
  left_join(model_meta) |>
  group_by(model_type) |> 
  summarise(n = n())  |> 
  mutate(site_id = 'TOOK', mean_skill = 3.25) |> 
  filter(model_type != 'null')

### 2024-09-03 supplementary_plots.Rmd
updates for review

```{r}
top_mods <- temperature_skill |> 
  group_by(model_id) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  ungroup() |> 
  filter(mean_crps > 0) |> 
  distinct(model_id)

@@

Plotting only dates where all sites are evaluated shows the same pattern as when all dates are considered. 
```{r overlapping-periods}
n_model_type <- temperature_skill |> 
  distinct(model_id) |> 
  left_join(model_meta) |>
  group_by(model_type) |> 
  summarise(n = n())  |> 
  mutate(site_id = 'TOOK', mean_skill = -3.25) |> 
  filter(model_type != 'null')

# top 10 average models
top_mods <- temperature_skill |> 
  group_by(model_id) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  ungroup() |> 
  slice_max(mean_crps, n = 10, na_rm = T) |> 
  distinct(model_id)

@@

A <- temperature_skill |> 
  left_join(model_meta, by = 'model_id') |>
  filter(model_type != 'null', 
         between(datetime, date_same_B$first_date, date_same_B$last_date)) |> 
  group_by(model_type, model_id, site_id) |> 
  summarise(mean_skill = mean(skill_crps, na.rm = T)) |>  
  ggplot(aes(x = mean_skill, 
             y = site_id, 
             colour = model_type)) +
  geom_vline(xintercept = 0, linetype = 'solid') +
  geom_label(data = n_model_type, aes(label = paste0('n = ', n)))+ 
  geom_boxplot() +
  facet_wrap(~model_type, scales = 'free_y',nrow = 2) +
  labs(y='Site', x = 'mean relative skill (°C)') +
  scale_colour_manual(values = cols_modeltype, name = 'Model Type') +
  theme_bw()

B <- temperature_skill |> 
  filter(model_id %in% top_mods$model_id, 
         between(datetime, date_same_B$first_date, date_same_B$last_date)) |> 
  group_by(site_id, horizon) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  mutate(site_id = factor(site_id, levels = lake_sites)) |> 
  ggplot(aes(x=horizon, 
             y = mean_crps, colour = site_id)) +
  geom_hline(yintercept = 0, colour = 'grey40') + 
  geom_line()  +  
  labs(y='mean relative skill (°C)', x = 'Horizon (days)') +
  scale_colour_brewer(palette = 'Dark2', name = 'Site') +
  theme_bw()

cowplot::plot_grid(A, B, nrow = 2, rel_heights = c(1.5, 1), align = 'v', axis = 'lr', labels = 'auto') |> 
  ggsave(filename = './Plots/FigureS4.png', width = 19, height = 16, units = 'cm')
```
```{r}
skill_summary_RW <- temperature_skill_RW |>
  group_by(model_id) |>
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |>
  ungroup() |>
  filter(mean_crps > 0) |>
  left_join(model_meta) |>
  mutate(manuscript_name = fct_reorder(manuscript_name, mean_crps),
         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no'),
         uses_AT = str_detect(NOAA_var, 'air_temperature')) |>
  ggplot(aes(x=mean_crps, y = manuscript_name, fill = model_type)) +    # Different pattern for each group
  geom_bar_pattern(aes(pattern = as.factor(uses_AT)),
                   stat = "identity",
                   alpha = 0.8,
                   pattern_fill = "black",
                   pattern_colour = "black",
                   pattern_spacing = 0.04,
                   pattern_alpha = 1,
                   colour = 'black') +
  scale_y_reordered() +
  labs(y='Model', x = 'Mean relative skill (°C)') +
  geom_vline(xintercept = 0) +
  scale_fill_manual(values = cols_modeltype) +
  scale_pattern_manual(values=c('none', 'stripe')) +
  scale_pattern_type_manual(values=c(NA, NA)) +
  theme_bw(base_size = 10) +
  guides(pattern = guide_legend(override.aes = list(fill = "white", pattern_spacing = 0.02),
                                title = 'Uses air temperature\nas a covariate?'),
         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model class')) +
    theme(legend.box = 'horizontal')

@@

ggpubr::ggarrange(ggarrange(NULL, as_ggplot(get_legend(skill_summary_RW)), NULL, NULL, widths =  c(0.2, 0.9,0.2, 1.1), nrow = 1),
                  ggarrange(skill_summary_RW_nolegend,  NULL, skill_horizon_RW, align = 'h', widths =  c(0.9,0.05, 1.1), nrow = 1), 
                  heights = c(0.5, 1),nrow = 2) |> 
  ggsave(filename = 'Plots/FigureSnew.png', height = 17, width = 18, units = 'cm')
```
```{r}
null_scores <- temperature_scores |> mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) %>% 
  filter(model_id %in% baselines$model_id, 
         horizon > 0, 
         horizon != 15,
         horizon <=30) |> 
  group_by(model_id, horizon) |> 
  summarise(mean_crps = mean(crps, na.rm = T)) %>% 
  mutate(model_id = factor(model_id, levels = c('climatology', 'persistenceRW'), labels = c("day-of-year", "persistence"))) %>% 
  ggplot(aes(x=horizon,y = mean_crps, colour = model_id)) + 
  geom_line() + 
  scale_x_continuous(breaks = seq(2,30, 2)) + 
  theme_bw() + 
  labs(x= 'Horizon (days)', y = 'Mean CRPS (degrees C)') + 
  scale_colour_viridis_d(name = 'Null Model', begin = 0.2, end=0.9) +
  theme(panel.grid.minor = element_blank())

@@

null_model_beats <- full_join(temp_skill_clim_summary, temp_skill_RW_summary, 
          by = join_by(model_id, horizon, manuscript_name)) %>% 
  mutate(beats = ifelse(mean_crpsDOY > 0 & mean_crpsRW > 0, 'both',
                        ifelse(mean_crpsDOY < 0 & mean_crpsRW > 0, 'RW',
                               ifelse(mean_crpsDOY > 0 & mean_crpsRW < 0, 'DOY', 'none')))) %>% 
  mutate(beats = factor(beats, levels = c("none", "RW", "DOY", "both"), labels = c("Neither", "Persistence", "Day-of-year", "Both"))) %>% 
  na.omit() %>% 
  ggplot(aes(x=horizon, y=manuscript_name, shape = beats, colour = beats)) + 
  geom_point(size = 5) +
  scale_shape_manual(values = c(0, 1, 4, 13), name = 'Null model that is out-performed') +
  scale_colour_manual(values = c("grey90", "#455BCDFF", "#E1DD37FF", "#46F884FF"), name = 'Null model that is out-performed') +
  scale_y_discrete(name = 'Model') +
  scale_x_continuous(name = 'Horizon (days)', breaks = c(1,10, 20, 30), expand = c(0.02,0)) +
  geom_hline(yintercept = seq(.5, 9.5, 1), linewidth = .2) +
  theme(legend.position = "bottom", 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        axis.ticks.y = element_blank())

>>>>

```{r}
top_mods <- temperature_skill |> 
  group_by(model_id) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  ungroup() |> 
  filter(mean_crps < 0) |> 
  distinct(model_id)

@@

Plotting only dates where all sites are evaluated shows the same pattern as when all dates are considered. 
```{r overlapping-periods}
n_model_type <- temperature_skill |> 
  distinct(model_id) |> 
  left_join(model_meta) |>
  group_by(model_type) |> 
  summarise(n = n())  |> 
  mutate(site_id = 'TOOK', mean_skill = 3.25) |> 
  filter(model_type != 'null')

# top 10 average models
top_mods <- temperature_skill |> 
  group_by(model_id) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  ungroup() |> 
  filter(mean_crps < 0) |> 
  distinct(model_id)

@@

A <- temperature_skill |>
  left_join(model_meta, by = 'model_id') |>
  filter(model_type != 'null',
         between(datetime, date_same_B$first_date, date_same_B$last_date)) |>
  group_by(model_type, model_id, site_id) |>
  summarise(mean_skill = mean(skill_crps, na.rm = T)) |>
  ggplot(aes(x = mean_skill,
             y = site_id,
             colour = model_type)) +
  geom_vline(xintercept = 0, linetype = 'solid') +
  geom_text(data = n_model_type, aes(label = paste0('n = ', n)))+
  geom_boxplot() +
  facet_wrap(~model_type, scales = 'free_y',nrow = 2) +
  labs(y='Site', x = 'mean relative skill (°C)') +
  scale_colour_manual(values = cols_modeltype, name = 'Model class') +
  theme_bw()

lake_order <- temperature_skill |> 
  filter(model_id %in% top_mods$model_id) |> 
  group_by(site_id, horizon) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) %>% 
  filter(horizon == 30) %>% 
  arrange(desc(mean_crps)) %>% 
  pull(site_id)

B <- temperature_skill |> 
  filter(model_id %in% top_mods$model_id, 
         between(datetime, date_same_B$first_date, date_same_B$last_date)) |> 
  group_by(site_id, horizon) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  mutate(site_id = factor(site_id, levels = lake_order)) |> 
  ggplot(aes(x=horizon, 
             y = mean_crps, colour = site_id)) +
  geom_hline(yintercept = 0, colour = 'grey40') + 
  geom_line()  +  
  labs(y='mean relative skill (°C)', x = 'Horizon (days)') +
  scale_colour_brewer(palette = 'Dark2', name = 'Site') +
  theme_bw()

cowplot::plot_grid(A, B, nrow = 2, rel_heights = c(1.5, 1), align = 'v', axis = 'lr', labels = 'auto') |> 
  ggsave(filename = './Plots/FigureS4.png', width = 19, height = 16, units = 'cm')
```
```{r}
skill_summary_RW <- temperature_skill_RW |>
  group_by(model_id) |>
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |>
  ungroup() |>
  filter(mean_crps < 0) |>
  left_join(model_meta) |>
  mutate(manuscript_name = fct_reorder(manuscript_name, mean_crps),
         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no'),
         uses_AT = str_detect(NOAA_var, 'air_temperature')) |>
  ggplot(aes(x=mean_crps, y = manuscript_name, fill = model_type)) +    # Different pattern for each group
  geom_bar_pattern(aes(pattern = as.factor(uses_AT)),
                   stat = "identity",
                   alpha = 0.8,
                   pattern_fill = "black",
                   pattern_colour = "black",
                   pattern_spacing = 0.04,
                   pattern_alpha = 1,
                   colour = 'black') +
  scale_y_reordered() +
  labs(y='Model', x = 'Mean relative skill (°C)') +
  geom_vline(xintercept = 0) +
  scale_fill_manual(values = cols_modeltype) +
  scale_pattern_manual(values=c('none', 'stripe')) +
  scale_pattern_type_manual(values=c(NA, NA)) +
  theme_bw(base_size = 10) +
  guides(pattern = guide_legend(override.aes = list(fill = "white", pattern_spacing = 0.02),
                                title = 'Uses air temperature\nas a covariate?'),
         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model class')) +
    theme(legend.box = 'horizontal')

@@

ggpubr::ggarrange(ggarrange(NULL, as_ggplot(get_legend(skill_summary_RW)), NULL, NULL, widths =  c(0.2, 0.9,0.2, 1.1), nrow = 1),
                  ggarrange(skill_summary_RW_nolegend,  NULL, skill_horizon_RW, align = 'h', widths =  c(0.9,0.05, 1.1), nrow = 1), 
                  heights = c(0.5, 1),nrow = 2) |> 
  ggsave(filename = 'Plots/FigureSnew.png', height = 17, width = 18, units = 'cm')
```
```{r}
null_scores <- temperature_scores |> mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) %>% 
  filter(model_id %in% baselines$model_id, 
         horizon > 0, 
         horizon != 15,
         horizon <=30) |> 
  group_by(model_id, horizon) |> 
  summarise(mean_crps = mean(crps, na.rm = T)) %>% 
  mutate(model_id = factor(model_id, levels = c('climatology', 'persistenceRW'), labels = c("day-of-year", "persistence"))) %>% 
  ggplot(aes(x=horizon,y = mean_crps, colour = model_id)) + 
  geom_line() + 
  scale_x_continuous(breaks = seq(2,30, 2)) + 
  theme_bw() + 
  labs(x= 'Horizon (days)', y = 'Mean CRPS (°C)') + 
  scale_colour_viridis_d(name = 'Null Model', begin = 0.2, end=0.9) +
  theme(panel.grid.minor = element_blank())

@@

null_model_beats <- full_join(temp_skill_clim_summary, temp_skill_RW_summary, 
          by = join_by(model_id, horizon, manuscript_name)) %>% 
  mutate(beats = ifelse(mean_crpsDOY < 0 & mean_crpsRW < 0, 'both',
                        ifelse(mean_crpsDOY > 0 & mean_crpsRW < 0, 'RW',
                               ifelse(mean_crpsDOY < 0 & mean_crpsRW > 0, 'DOY', 'none')))) %>% 
  mutate(beats = factor(beats, levels = c("none", "RW", "DOY", "both"), labels = c("Neither", "Persistence", "Day-of-year", "Both"))) %>% 
  na.omit() %>% 
  ggplot(aes(x=horizon, y=manuscript_name, shape = beats, colour = beats)) + 
  geom_point(size = 5) +
  scale_shape_manual(values = c(0, 1, 4, 13), name = 'Null model that is out-performed') +
  scale_colour_manual(values = c("grey90", "#455BCDFF", "#E1DD37FF", "#46F884FF"), name = 'Null model that is out-performed') +
  scale_y_discrete(name = 'Model') +
  scale_x_continuous(name = 'Horizon (days)', breaks = c(1,10, 20, 30), expand = c(0.02,0)) +
  geom_hline(yintercept = seq(.5, 9.5, 1), linewidth = .2) +
  theme(legend.position = "bottom", 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        axis.ticks.y = element_blank())
